---
title: "SMS Spam Classification"
author: "Deepal DSilva"
date: "July 10, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Filtering SMS spam messages using various classification algorithms such as Naive Bayes, SVM, Decision Trees & Random 

```{r}
setwd("C:/Users/dsilv/Desktop/Learning/Data Science/Data-Analysis-Projects-In-R/Spam Text Classification")
```


Import Libraries 
```{r}
library(tm)
library(SnowballC)
library(wordcloud)
library(RColorBrewer) 
library(randomForest)  #For Random Forests
library(rpart)         #For Decision Trees
library(e1071)         #For Naive Bayes & SVM
library(caret)         #For the confusion matrix
```


Import data

```{r}
sms_raw <- read.csv("spam.csv")
```

1. Data Exploration & Visualization

```{r}
head(sms_raw)

```
    
We see that there are only 5 columns: first indicating if it a spam/ham(non-spam) message and the second with the actual message. The last 3 do not have a lot of data.
 We are going to remove the last 3 columns from the dataset and also rename the column headers.

```{r}

sms_raw <- sms_raw[, 1:2]

colnames(sms_raw) <- c("Tag", "Msg")

str(sms_raw)

table(sms_raw$Tag)

prop.table(table(sms_raw$Tag))


```

We see around 87% of the messages are valid and 13% are spam.

Let's visualize the spam and ham sms messages using wordclouds.

```{r}
spam <- subset(sms_raw, Tag == "spam")
wordcloud(spam$Msg, max.words = 60, colors = brewer.pal(7, "Paired"), random.order = FALSE)


ham <- subset(sms_raw, Tag == "ham")
wordcloud(ham$Msg, max.words = 60, colors = brewer.pal(7, "Paired"), random.order = FALSE)
```


2. Data Preprocessing
```{r}

#Data Cleansing
sms_corpus <- VCorpus(VectorSource(sms_raw$Msg))

sms_dtm <- DocumentTermMatrix(sms_corpus, control = 
                                 list(tolower = TRUE,
                                      removeNumbers = TRUE,
                                      stopwords = TRUE,
                                      removePunctuation = TRUE,
                                      stemming = TRUE))

dim(sms_dtm)

#Prepare the training and test dataset 80% and 20%. As the data is already randomly sorted we can split it directly.

#Training set
sms_dtm_train <- sms_dtm[1:4457, ]

#Test set
sms_dtm_test <- sms_dtm[4458:5572, ]

#Training Label
sms_train_labels <- sms_raw[1:4457, ]$Tag

#Test Label
sms_test_labels <- sms_raw[4458:5572, ]$Tag

#Proportion for train labels
prop.table(table(sms_train_labels))

#Proportion for test labels
prop.table(table(sms_test_labels))

```

#Naive Bayes
Creating indicator features

We extract the most frequent words in the texts to reduce the number of features in the Naive Bayes algorithm

```{r}
threshold <- 0.1

min_freq = round(sms_dtm$nrow*(threshold/100),0)

min_freq

# Create vector of most frequent words
freq_words <- findFreqTerms(x = sms_dtm, lowfreq = min_freq)

str(freq_words)
```

We now have a vector of 1260 words that are most frequent.

```{r}
#filter the DTM sparse matrix to only contain words that are most frequent to reduce features in the DTM

sms_dtm_freq_train <- sms_dtm_train[ , freq_words]
sms_dtm_freq_test <- sms_dtm_test[ , freq_words]

dim(sms_dtm_freq_train)


```

#Convert numeric values into categorical values

Since Naive Bayes trains on categorical data, the numerical data must be converted to categorical data. We will convert the numeric features by creating a function that converts any non-zero positive value to "Yes" and all zero values to "No" to indicate whether a specific term is present in the document.

```{r}
convert_values <- function(x) {
  x <- ifelse(x > 0, "Yes", "No")
}

sms_train <- apply(sms_dtm_freq_train, MARGIN = 2,
                   convert_values)
sms_test <- apply(sms_dtm_freq_test, MARGIN = 2,
                  convert_values)

str(sms_train)
```

3. ML Model Fitting

Naive Bayes
```{r}
sms_classifier <- naiveBayes(sms_train, sms_train_labels)

sms_test_pred <- predict(sms_classifier, sms_test)


```


4. Evaluating the model

# Create confusion matrix
```{r}
confusionMatrix(data = sms_test_pred, reference = sms_test_labels, positive = "spam", dnn = c("Prediction", "Actual"))
```

