---
title: "Data Scientist Survey"
author: "Deepal DSilva"
date: "August 23, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction


## Installing Required Packages

```{r Import PACKAGES, message=FALSE, warning=FALSE}
setwd("C:/Users/dsilv/Desktop/Learning/Data Science/Data-Analysis-Projects-In-R/Data Scientist Survey")

library(tidyverse)
library(stringr)
library(tm)
library(rlang)
library(splitstackshape)
```

## Loading Data
The data comes in 4 files:

* Multiple choice Data
* Free form responses
* Questions asked
* Currency conversion rates (for the salary)

```{r Data import, message=FALSE, warning=FALSE}
# Import multiple choice responses
rawMCdata <- read.csv('multipleChoiceResponses.csv', stringsAsFactors = TRUE, header = TRUE)

# Import freeform responses
rawFFRdata <- read.csv('freeformResponses.csv', stringsAsFactors = FALSE, header = TRUE)

# Import questions asked
schema <- read.csv('schema.csv', stringsAsFactors = FALSE, header = TRUE)

# Import currency conversion rates
conversionRates <- read.csv('conversionRates.csv', header = TRUE)

```

Functions to process Multiple Choice survey response sheet

```{r}
# A function to analyze responses which are a single choice
singleChoice = function(question, filteredData = rawMCdata){
  

  filteredData %>% 
    # Remove any rows where the respondent didn't answer the question
    drop_na(question)  %>% 
    # Group by the responses to the question
    group_by(!! (sym(question))) %>% 
    # Count how many respondents selected each option
    summarise(count = n()) %>% 
    # Calculate what percent of respondents selected each option
    mutate(percent = round((count / sum(count)) * 100, digits=2)) %>% 
    # Arrange the counts in descending order
    arrange(desc(count))
  
}

```


Gender distribution
```{r}
gender <- singleChoice("GenderSelect") #%>%
            filter(GenderSelect == "Male" | GenderSelect == "Female" )


ggplot(gender,aes(x= reorder(GenderSelect, -percent),
                  y= percent, fill=GenderSelect)) + geom_bar(stat="identity") +
labs(x="Gender",y="Percent",title="Distribution of Gender") +  geom_text(aes(label=percent),color="black") + 
  theme(legend.position="none") + 
  theme(plot.title = element_text(hjust = 0.5))

#____________

#genderlevels <- c("Male","Female","A different identity","Non-binary, genderqueer, or gender non-conforming")
#gender$GenderSelect = factor(gender$GenderSelect,levels=genderlevels) %>% droplevels() %>% filter(!is.na(GenderSelect))





```
0.48% identify as "Non-binary, genderqueer, or gender on-conforming", and .98% identify as "A different identity" . :et's remove this from our analysis since this is going to vary greatly across countries in our furher analysis.

Age Distribution

```{r}
rawMCdata$Age <- as.numeric(as.character(rawMCdata$Age))

agedf <- rawMCdata %>% 
  filter(!Age == "") %>% 
  select(Age)

ggplot(agedf, aes(Age), binwidth=2) + 
  geom_histogram() + 
  xlab("Age") + 
  ylab("Number of Respondents")


```
The majority of Kaggle users are young adults (early 20's to 30's). We also see some entries at 0 & 100 years that are likely not accurate. 

Lets compare our age & gender among the top 10 countries
### Data Exploration
### Demographics - Age, Gender, Location, Age/gender, Location/gender, Education

### Employment - Type of Industry, Size of company, Tools used a work, DS methods at wrk, Maths level reqd at work, remote work, sharing code at work, future learning, recommended language, Annual Salary, Job satisfaction, Promotions/Salary increase

### Career background

### Breaking into DS - Resources, Podcast/Blog etc, Time spent learning DS, Important skills, Online platforms, Time spent per week on studying, Proving/Showcase your knowledge, where to look for job opps, Area & Technique of Competence